Legion: Expressing Locality and Independence with Logical Regions

### 2012-05-05 doc/sc2012/related.tex

Legion began as an outgrowth of Sequoia, a locality-aware programming language capable
of expressing computations in deep memory hierarchies
\cite{Fatahalian06}.  Sequoia is a special case of the Legion
programming model in which only arrays can be recursively partitioned,
all access is exclusive, there is a static mapping of tasks and data on to the machine (though
extensions to Sequoia make this mapping more dynamic \cite{BCSA11}) and, most fundamentally, the decomposition of tasks and the decomposition
of data is one-to-one.  Legion generalizes the Sequoia model by allowing for
partitioning of pointer data structures through regions, enabling dynamic mappings through
the mapper interface, and allowing different coherence
properties.  Legion's decoupling of the task tree from the region tree in Legion leads directly to the scheduling problem solved by our software out-of-order processor for tasks with region arguments.

@@

X10 is another parallel programming environment designed to operate on distributed
memory machines \cite{X1005}.  X10's {\em places} enable programmers to talk about where
to place both data and tasks in the machine.  However, once data and tasks have
been placed they are fixed, which mandates that data movement be explicitly managed by
user level code or implicitly by the compiler or runtime \cite{X1008}.  Recently X10 has introduced
the concept of a region into the compiler's intermediate representation \cite{X1011}.
Unlike Legion, regions in X10 are not visible to the programmer but are inferred
from high level arrays through static analysis.  X10 provides support for clusters of GPUs
\cite{X10GPU}, but requires the programmer to write all code managing data movement
through both the cluster and GPU memory hierarchies.

@@

SPMD languages such as Titanium \cite{TIT98} and UPC \cite{UPC99} have
mechanisms for describing array partitions in distributed memories.
However, the partition operations supported only operate on two-level
memory hierarchies consisting of local and global memory. 
% Neither
%Titanium nor UPC currently support GPUs as part of their language
%standard \cite{TITANIUMSTANDARD} \cite{UPCSTANDARD}.  
Part of Legion's
low-level runtime system is constructed using UPC's underlying GASNet
runtime system \cite{GASNET07}.

>>>>

Legion began as an outgrowth of Sequoia, a locality-aware programming language capable
of expressing computations in deep memory hierarchies
\cite{Fatahalian06}.  Sequoia is a special case of the Legion
programming model in which only arrays can be recursively partitioned,
all access is exclusive, there is a static mapping of tasks and data (though
extensions to Sequoia make this mapping more dynamic \cite{BCSA11}) and, most fundamentally, the decomposition of tasks and the decomposition
of data is one-to-one.  Legion generalizes the Sequoia model by allowing for
partitioning of pointer data structures through regions, enabling dynamic mappings through
the mapper interface, and allowing different coherence
properties.  Legion's decoupling of the task tree from the region tree leads directly to the scheduling problem solved by our software out-of-order processor for tasks with region arguments.

@@

X10 is another parallel programming language designed to operate on
distributed memory machines \cite{X1005}.  X10's {\em places} enable
programmers to talk about where to place both data and tasks.
However, once data and tasks have been placed they are fixed, which
mandates that data movement be explicitly managed by user level code
or implicitly by the compiler or runtime \cite{X1008}.  Recently X10
has introduced regions into the compiler's
intermediate representation \cite{X1011}.  Unlike Legion, regions in
X10 are not visible to the programmer but are inferred from high level
arrays through static analysis.  X10 provides support for clusters of
GPUs \cite{X10GPU}, but requires the programmer to write all code
managing data movement through both the cluster and GPU memory
hierarchies.

@@

SPMD languages such as Titanium \cite{TIT98} and UPC \cite{UPC99} have
mechanisms for describing array partitions in distributed memories.
However, the partition operations supported only operate on two-level
memory hierarchies consisting of local and global memory. 
% Neither
%Titanium nor UPC currently support GPUs as part of their language
%standard \cite{TITANIUMSTANDARD} \cite{UPCSTANDARD}.  
Part of Legion's
low-level runtime system is constructed using UPC's GASNet
runtime system \cite{GASNET07}.
